Incentivizing Exploration by Heterogeneous Users

### 2018-06-06 Lemmas.tex

\begin{rlemma}{Lemma}{\ref{lem:phase-length}}
For any $s\geq 3$, the expected length of phase $s$ is at most
$\ARMNUM \cdot \log(s)$ time steps.
\pfedit{
Moreover, for any set of types $A$, the expected number of times that 
an agent with a type in $A$ appears in a phase $s$ is at most 
$f(A) \cdot \ARMNUM \cdot \log(s)$,
where $f(A) := \Prob[\AgV \sim \AgentDist]{\ArmV \in A}$.
}
\end{rlemma}

\begin{proof}
\pfedit{
We show the second claim for general $A$, which implies the first claim by taking $A$ to be the support of $\AgentDist$.

Fix a phase $s$ and let $S_t$ be the set of arms that have been pulled at most $s$ times by time $t$.  Let $\tau_m$ be the maximum of $t_s$ (the start of phase $s$) and the time $t$ when $S_t$ first contains $m$ elements.  The phase ends when $S_t$ contains no elements, i.e., $\tau_0 = t_{s+1}$.
We consider the expected number of pulls during times $[\tau_m, \tau_{m-1})$ made by agents whose types are in $A$.

Consider a counter starting from $0$ at time $\tau_m$ that increments each time an agent arrives with a type in $A$, and that is stopped the first time an arm in $S_t$ is pulled.
For any stopping time $t \in [\tau_m, \tau_{m-1})$, 
the conditional probability that the counter increments
given the history of pulls, payments, and observations at time $t$ 
is $f(A)$.
The conditional probability that the counter is stopped by that agent's pull is at least $1/\log(s)$.  To see this, consider two cases.  In the first case, there is at least one payment eligible arm, in which case the principal incentives an arm in $S_t$.  In the second case, there are no payment eligible arms.  
Then, by definition of payment-eligibility,
each arm $i \in S_t$ has conditional probability at least $1/\log(s)$ of being pulled.

The expected value of this counter when it is stopped is bounded above by the expected stopped value of another counter whose conditional probability of being stopped is exactly $1/\log(s)$.  (This can be shown more formally via a coupling argument.)
Let $V_\ell$ be the conditional expectation of this second counter's value when it is stopped,
given that its current value is $\ell$ and it has not been stopped.  Observe that $V_\ell = \ell + V_0$.
Enumerating outcomes (either the counter increments or not; either it stops or it continues) provides $V_0 = f(A)\frac1{\log(s)} + f(A) (1-\frac1{\log(s)}) V_1 + (1-f(A))(1-\frac1{\log(s)}) V_0$.
Using $V_1 = 1 + V_0$ simplifies this to 
$V_0 = f(A) + (1-\frac1{\log(s)}) V_0$.
Noting $V_0$ is finite and solving for $V_0$ provides
$V_0 = f(A) \log(s)$.
Thus, the conditional expected number of pulls (given the history at time $\tau_m$) by an agent with a type in $A$ during times in $[\tau_m, \tau_{m-1})$ is no more than $f(A) \log(s)$.

We may write the total number of pulls by agents with types in $A$ during phase $s$ as a sum of this quantity over $m$ ranging from $|S_{t_s}|$ down to 1.  Noting that $|S_{t_s}| \le N$ and taking the expectation of this sum shows the claim. 
}
\end{proof}

>>>>

\begin{rlemma}{Lemma}{\ref{lem:phase-length}}
For any $s\geq 3$, the expected length of phase $s$ is at most
$\ARMNUM \cdot \log(s)$ time steps.

More generally, for any set of types $A$, the expected number of times that 
an agent with a type in $A$ appears in a phase $s$ is at most 
$f(A) \cdot \ARMNUM \cdot \log(s)$,
where $f(A) := \Prob[\AgV \sim \AgentDist]{\dkedit{\AgV} \in A}$.
\end{rlemma}

\begin{proof}
We show the second claim for general $A$,
which implies the first claim by taking $A$ to be the support of $\AgentDist$.

Fix a phase $s$, and let $S_t$ be the set of arms that have been
pulled at most $s$ times by time $t$.
Let $\tau_m$ be the maximum of $t_s$ (the start of phase $s$)
and the time $t$ when $S_t$ first contains \dkedit{(at most)} $m$ elements.
The phase ends when $S_t$ contains no elements, i.e., $\tau_0 = t_{s+1}$.
We consider the expected number of pulls during times
$\dkedit{t \in} [\tau_m, \tau_{m-1})$ made by agents whose types are in $A$.

Consider a counter starting from $0$ at time $\tau_m$ that increments
each time an agent arrives with a type in $A$,
and that is stopped the first time an arm in $S_t$ is pulled.
For any stopping time $t \in [\tau_m, \tau_{m-1})$, 
the conditional probability that the counter increments,
given the history of pulls, payments, and observations at time $t$,
is $f(A)$.
The conditional probability that the counter is stopped by that
agent's pull is at least $1/\log(s)$.
To see this, consider two cases.
In the first case, there is at least one payment-eligible arm,
in which case the principal incentivizes an arm in $S_t$.
In the second case, there are no payment-eligible arms.
Then, by definition of payment-eligibility,
each arm $i \in S_t$ has conditional probability at least $1/\log(s)$ of being pulled.

The expected value of this counter when it is stopped is bounded above
by the expected stopped value of another counter whose conditional
probability of being stopped is exactly $1/\log(s)$.
(This can be shown more formally via a coupling argument.) 
Let $V_\ell$ be the conditional expectation of this second counter's value when it is stopped,
given that its current value is $\ell$ and it has not been stopped.
Observe that $V_\ell = \ell + V_0$.
Enumerating outcomes
(either the counter increments or not; either it stops or it continues)
gives us the recurrence relation
$V_0 = f(A) \frac{1}{\log(s)} + f(A) (1-\frac{1}{\log(s)}) V_1
   + (1-f(A))(1-\frac{1}{\log(s)}) V_0$.
Using that $V_1 = 1 + V_0$, this equation simplifies to 
$V_0 = f(A) + (1-\frac{1}{\log(s)}) V_0$.
Noting that $V_0$ is finite and solving for $V_0$ gives us that
$V_0 = f(A) \log(s)$.
Thus, the conditional expected number of pulls (given the history at time $\tau_m$)
by an agent with a type in $A$ during times in $[\tau_m, \tau_{m-1})$
is no more than $f(A) \log(s)$.

We may write the total number of pulls by agents with types in $A$
during phase $s$ as a sum of this quantity over $m$ ranging from
$|S_{t_s}|$ down to 1.
Noting that $|S_{t_s}| \leq \ARMNUM$ and taking the expectation of
this sum completes the proof. 
\end{proof}

### 2018-06-06 algorithm_general.tex

We begin by bounding the length of any phase, which will support bounding
the regret of early rounds
(before tail bounds have kicked in). The proof of Lemma~\ref{lem:phase-length} and all other proofs missing in the main paper may be found in the appendix.

\begin{lemma} \label{lem:phase-length}
For any $s\geq 3$, the expected length of phase $s$ is at most
$\ARMNUM \cdot \log(s)$ time steps.
\end{lemma}

>>>>

We begin by bounding the length of any phase,
\dkedit{and more generally, the number of arm pulls by any given set
  of agents.
The bound of Lemma~\ref{lem:phase-length}} will support bounding the
regret of early rounds (before tail bounds have kicked in).
The proof of Lemma~\ref{lem:phase-length} and all other proofs missing
in the main paper may be found in the appendix.

\begin{lemma} \label{lem:phase-length}
For any $s\geq 3$, the expected length of phase $s$ is at most
$\ARMNUM \cdot \log(s)$ time steps.

More generally, for any set of types $A$, the expected number of times that 
an agent with a type in $A$ appears in a phase $s$ is at most 
$f(A) \cdot \ARMNUM \cdot \log(s)$,
where $f(A) := \Prob[\AgV \sim \AgentDist]{\dkedit{\AgV} \in A}$.
\end{lemma}

%For any $s\geq 3$, the expected length of phase $s$ is at most
%$\ARMNUM \cdot \log(s)$ time steps.

### 2018-06-06 algorithm_regret.tex

For the remainder of the proof, we consider phases $s$
with $\EarlyCut{s} \leq \TieCutoff$.
\pfedit{
We first consider agents \AgV incurring positive regret less than \EarlyCut{s}.
Let $A$ be the set of types satisfying 
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq \EarlyCut{s}$.
By Assumption~\ref{A1}, and because $\EarlyCut{s} \leq \TieCutoff$,
the total measure of $A$
is $\AlmostTied{\EarlyCut{s}} \leq \TieDensity \cdot \EarlyCut{s}$,
Then, by Lemma~\ref{lem:phase-length},
the expected number of pulls in phase $s$ by agents in $A$  
bounded above by $\TieDensity\ \EarlyCut{s} \ARMNUM \log(s)$, and the total expected number of pulls by these agents by $\sum_{s=1}^T \TieDensity \ \EarlyCut[2]{s} \ARMNUM \log(s)$.
Consider any agent $\AgV$ incurring positive regret less than $\EarlyCut{s}$.
This agent's type must be in $A$.
Moreover, this agent's regret is bounded above by $\EarlyCut{s}$.
Thus, the expected total regret from such agents is at most
\begin{equation}
\sum_{s=1}^{T} \TieDensity \cdot \EarlyCut[2]{s} \ARMNUM \log(s).
\label{equ:small_regret_bound}
\end{equation}
}

We next bound the regret incurred in time steps with large regret.
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > \pfedit{t_{s+1}-t_{s}}$.
By Lemma~\ref{lem:right-choice},
under \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}}, no agent at time $t$ will incur
regret more than \EarlyCut{s}.

@@

In the low-probability event, we simply 
bound the maximum regret by \MAXR.
Now, applying Lemma~\ref{lem:phase-length},
the total expected regret from large-regret steps is at most
$\sum_{s = 1}^T \frac{24 \ARMNUM d}{T^2} \cdot \MAXR \cdot \ARMNUM \log(s)
\leq 12 \ARMNUM^2 d \MAXR.$
%\pfcomment{I deleted the left-hand side of the above equation because it was using random terms in the summand, and actually wasn't true.  What was intended was an expectation outside of the sum, but I thought it simpler just to delete that term.}

Adding all four types of regret terms,
and substituting 
$\EarlyCut[2]{s} = \frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most
$
O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \MAXR \log^2(T)
+ \ARMNUM^2 \MAXR d
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} \EarlyCut[2]{s} \log(s) \right)\\
= 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \TieDensity \cdot \log^3(T) \right)$.\QED
\end{emptyextraproof}

>>>>

For the remainder of the proof, we consider phases $s$
with $\EarlyCut{s} \leq \TieCutoff$.
We first consider agents \AgV incurring positive regret less than \EarlyCut{s}.
Let $A_s$ be the set of types satisfying 
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq \EarlyCut{s}$.
By Assumption~\ref{A1}, and because $\EarlyCut{s} \leq \TieCutoff$,
the total measure of $A_s$
is $\AlmostTied{\EarlyCut{s}} \leq \TieDensity \cdot \EarlyCut{s}$.
Then, by Lemma~\ref{lem:phase-length},
the expected number of pulls in phase $s$ by agents in $A_s$  
is bounded above by $\TieDensity \EarlyCut{s} \cdot \ARMNUM \log(s)$.
\dkdelete{and the total expected number of pulls by these agents by
$\sum_{s=1}^T \TieDensity \EarlyCut[2]{s} \ARMNUM \log(s)$.
Consider}
Any agent \AgV incurring positive regret less than $\EarlyCut{s}$
must have a type in $A_s$,
\dkdelete{Moreover, this agent's regret is bounded above by $\EarlyCut{s}$.}
so the expected total regret from such agents,
\dkedit{summed over all phases,} is at most
\begin{equation}
\sum_{s=1}^{T} \TieDensity \EarlyCut[2]{s} \cdot \ARMNUM \log(s).
\label{equ:small_regret_bound}
\end{equation}

We next bound the regret incurred in time steps with large regret.
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > t_{s+1}-t_{s}$.
By Lemma~\ref{lem:right-choice},
under \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}}, no agent at time $t$ will incur
regret more than \EarlyCut{s}.

@@

In the low-probability event, we simply 
bound the maximum regret by \MAXR.
Summing over the $T$ time periods,
the total expected regret from large-regret steps is at most
$24 \ARMNUM d \MAXR / T$.

Adding all four types of regret terms,
and substituting 
$\EarlyCut[2]{s} = \frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most
$
O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \MAXR \log^2(T)
+ \frac{\ARMNUM \MAXR d}{T}
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} \EarlyCut[2]{s} \log(s) \right)\\
= 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \TieDensity \cdot \log^3(T) \right)$.\QED
\end{emptyextraproof}

### 2018-06-06 problem_setting.tex

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms are identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
and the principal.
When arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero independent \pfedit{continuous} sub-Gaussian%
\footnote{For simplicity of notation, we assume that the variance \pfedit{proxy}
  $\sigma^2$ is uniform across time steps.
  \pfedit{The} analysis extends \pfedit{easily when the variance proxy changes over time}.}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.

>>>>

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms are identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
and the principal.
When arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero
independent continuous sub-Gaussian%
\footnote{For simplicity of notation, we assume that the variance proxy
  $\sigma^2$ is uniform across time steps.
  The analysis extends easily when the variance proxy changes over time.}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.
